Estimating the number of migrants using between-population sibling pairs

### 2021-05-21 latex/ms.tex
start #4
add simple POP-model
define additional estimators including those for migration rate

@@

\subsection{Basic model}

まず最初に，最もシンプルなモデルを考える。$t$年にサブ集団１にいる$N_{1,t}$個体の母親を考えよう。$t$年に子供を産んだ後，$N_{1,t}$個体のうち$M_{1\to2}$個体がサブ集団２に移動し，$t+1$年にサブ集団２にいる母個体$N_{2,t+1}$の一部として子供を産むとする。この時，サブ集団１から$t$年に産まれた子供とサブ集団２から$t+1$年に産まれた子供が，母を共有する兄弟(Maternal half Sibling, MHS)である確率$\pi_{\rm b,1\to2}$を考える。どの母親も同じような繁殖ポテンシャルを持つと仮定すると，移動する母親の子供をサブ集団１でサンプルする確率は$M_{1\to2}/N_{1,t}$となる。同様に，移動する母親の子供をサブ集団２においてサンプルする確率は$M_{1\to2}/N_{2,t+1}$である。それぞれでサンプルした子供の母親が同一である確率は$1/M_{1\to2}$であることより，サンプルした２個体がMHSペアとなる確率が以下のように得られる。
\begin{align}
\pi_{\rm b,1\to2} &= \frac{M_{1\to2}}{N_{1,t}} \frac{M_{1\to2}}{N_{2,t+1}} \frac{1}{M_{1\to2}} \nonumber\\
&= \frac{M_{1\to2}}{N_{1,t}N_{2,t+1}}
\label{1-1}
\end{align}
サブ集団１からサブ集団２に移動する割合$m_{1\to2}$を導入することで，上式は
\begin{align}
\pi_{\rm b,1\to2} &= \frac{m_{1\to2}}{N_{2,t+1}}
\label{1-2}
\end{align}
と書ける。$M_{1\to2} = m_{1\to2}N_{1,t}$とした。$m_{1\to2}$には移動だけでなく$t+1$年の産卵までの生存確率が含まれている点に注意。

@@

@@

\section{Application}\label{sec3}

ちなみに、同一年内に同じサブ集団内で見つかるMHSペアについては、有効集団サイズの情報（正確には有効繁殖サイズ、Effective breeding size）は持っているが、センサスサイズに関するダイレクトな情報は持っていない点に注意。

\section{Discussion}\label{sec4}

>>>>

\subsection{Assumptions}

@@

\subsection{HSP-based model}

\begin{align}
\pi_{\rm HS,bet} &= \frac{2M}{N_{1}} \frac{2M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{4M}{N_{1}N_{2}}
\label{HS-1}
\end{align}

\begin{align}
\mathbb{E}[H_{\rm HS, bet}] &= \pi_{\rm HS,bet} n_{\rm O, 1} n_{\rm O, 2} \nonumber\\
&= \frac{ 4 n_{\rm O, 1} n_{\rm O, 2} M } { N_{1} N_{2} } \\
&= \frac{ 4 n_{\rm O, 1} n_{\rm O, 2} m } { N_{2} }
\label{HS-2}
\end{align}

\begin{align}
\widehat{M_1} &= \frac{N_{1} N_{2} \widetilde{H}_{\rm HS, bet} } {4 n_{\rm O, 1} n_{\rm O, 2} }
\label{HS-3}
\end{align}

\begin{align}
\widehat{m_1} &= \frac{N_{2} \widetilde{H}_{\rm HS, bet} } {4 n_{\rm O, 1} n_{\rm O, 2} }
\label{HS-3}
\end{align}

より一般的な説明は付録に載せた。

\subsection{POP-based model}

\begin{align}
\pi_{\rm PO,bet} &= \frac{2M}{N_{1}} \frac{M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{2M}{N_{1}N_{2}}
\label{PO-1}
\end{align}

\begin{align}
\mathbb{E}[H_{\rm PO, bet}] &= \pi_{\rm PO,bet} n_{\rm O, 1} n_{\rm P, 2} \nonumber\\
&= \frac{ 2 n_{\rm O, 1} n_{\rm P, 2} M } { N_{1} N_{2} }
\label{PO-2}
\end{align}

\begin{align}
\widehat{M_2} &= \frac{N_{1} N_{2} \widetilde{H}_{\rm PO,bet} } {2 n_{\rm O, 1} n_{\rm P, 2} }
\label{PO-3}
\end{align}

\section{Applications}\label{sec3}

\subsection{Parent number can be estimated by PO pairs found within a subpopulation}

\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) }
\label{AP-1}
\end{align}

\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) }
\label{AP-2}
\end{align}

サンプル数の少なさを補えるかも。

\begin{align}
\widehat{N_2} &= \frac{ 2 n_{\rm O, 2} n_{\rm P, 2} + 1 }{\widetilde{H}_{\rm PO, 2} + 1} 
\label{AP-3}
\end{align}
subpop2の親子もわかるはず。migration rateを推定すると考えると、$N_2$を推定値で置き換えて

\begin{align}
\widehat{m_3} &=  \frac{ \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) \left(2 n_{\rm O, 2} n_{\rm P, 2} + 1\right)} {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}
\label{AP-4}
\end{align}

\begin{align}
\widehat{N_1} &= \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} + 1 }{\widetilde{H}_{\rm PO, 1} + 1} 
\label{AP-5}
\end{align}

\begin{align}
\widehat{M_5} &=  \frac{ \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) \left(2 n_{\rm O, 1} n_{\rm P, 1} + 1\right) \left(2 n_{\rm O, 2} n_{\rm P, 2} + 1\right)} {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 1} + 1\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}
\label{AP-6}
\end{align}

\begin{comment}%%%%%%%%%%%%%%%
まず最初に，最もシンプルなモデルを考える。$t$年にサブ集団１にいる$N_{1,t}$個体の母親を考えよう。$t$年に子供を産んだ後，$N_{1,t}$個体のうち$M_{1\to2}$個体がサブ集団２に移動し，$t+1$年にサブ集団２にいる母個体$N_{2,t+1}$の一部として子供を産むとする。この時，サブ集団１から$t$年に産まれた子供とサブ集団２から$t+1$年に産まれた子供が，母を共有する兄弟(Maternal half Sibling, MHS)である確率$\pi_{\rm b,1\to2}$を考える。どの母親も同じような繁殖ポテンシャルを持つと仮定すると，移動する母親の子供をサブ集団１でサンプルする確率は$M_{1\to2}/N_{1,t}$となる。同様に，移動する母親の子供をサブ集団２においてサンプルする確率は$M_{1\to2}/N_{2,t+1}$である。それぞれでサンプルした子供の母親が同一である確率は$1/M_{1\to2}$であることより，サンプルした２個体がMHSペアとなる確率が以下のように得られる。
\begin{align}
\pi_{\rm b,1\to2} &= \frac{M_{1\to2}}{N_{1,t}} \frac{M_{1\to2}}{N_{2,t+1}} \frac{1}{M_{1\to2}} \nonumber\\
&= \frac{M_{1\to2}}{N_{1,t}N_{2,t+1}}
\label{1-1}
\end{align}
サブ集団１からサブ集団２に移動する割合$m_{1\to2}$を導入することで，上式は
\begin{align}
\pi_{\rm b,1\to2} &= \frac{m_{1\to2}}{N_{2,t+1}}
\label{1-2}
\end{align}
と書ける。$M_{1\to2} = m_{1\to2}N_{1,t}$とした。$m_{1\to2}$には移動だけでなく$t+1$年の産卵までの生存確率が含まれている点に注意。

@@

ちなみに、同一年内に同じサブ集団内で見つかるMHSペアについては、有効集団サイズの情報（正確には有効繁殖サイズ、Effective breeding size）は持っているが、センサスサイズに関するダイレクトな情報は持っていない点に注意。
\end{comment}%%%%%%%%%%%%%%%

\section{Results and Discussion}\label{sec4}

\begin{enumerate}
  \item $M$ではなく$m$の推定であれば、N2だけ分かれば良い
  \item 
  \item サンプルされる当歳魚（０歳）は，必ずサンプルされたサブ集団で産まれる
  \item 当歳魚かどうかの判断は１００％可能
  \item サンプル間の近親判別では，１００％の成功率で半兄弟ペアか否かがなされる
\end{enumerate}

\clearpage

\section*{APPENDIX A}
\setcounter{equation}{0}

\section*{Distribution of the number of surviving offspring per mother and its variance}

\renewcommand{\theequation}{A\arabic{equation}}

@@

Let $k_i$ denote the total number of surviving offspring from a mother  $i$ ($i= 1, 2, \dots, N_{\rm m}$) at time of sampling. Following \cite{Akita_2019} and giving the expected number of the surviving offspring per mother at time of sampling, $\lambda_i$ ($>0$), $k_i$ is assumed to follow a negative binomial distribution through a conventional parametrization:
\begin{equation}
{\rm Pr}[ k_i | \lambda_i ] = \frac{\Gamma[k_i+\phi]}{k_i! \Gamma[\phi]} \left( \frac{\lambda_i}{\phi+\lambda_i} \right)^{k_i} \left( \frac{\phi}{\phi+\lambda_i} \right)^{\phi}, 
\label{NB}
\end{equation}
where $\phi$ ($>0$) denotes the overdispersion parameter describing the degree of nonparental variation. In the current work, $\phi$ is assumed to be constant across mothers, whereas the expected number of surviving offspring ($\lambda_i$) varies across mothers. The mean and variance of this distribution are denoted by $\lambda_i$ and $\lambda_i + \lambda_i^2/\phi$, respectively. In the limit of infinite $\phi$, this distribution becomes a Poisson distribution, which is given by ${\rm Pr}[ k_i | \lambda_i ]=\lambda_i^{k_i} \mathrm{e}^{-\lambda_i}/(k_i!)$. We assume $\lambda_i$ to be independent and identically distributed with the density function $f(\lambda)$, producing a parental variation. The shape of the density function is often complex, but may be described by information from the population, for example, the mother's weight composition in the population. The specific form of $f(\lambda)$ is given in {\bf Appendix B} and is used for running an individual-based model.

The variance of the offspring number can be given by 
\begin{eqnarray}
\mathbb{V}[k] &=& \mathbb{E}[ \mathbb{V}[k|\lambda] ] + \mathbb{V}[ \mathbb{E}[k|\lambda] ] \nonumber\\
&=& \mathbb{E}[ \lambda + \lambda^2 / \phi ] + \mathbb{V}[ \lambda ].
\label{V_k}
\end{eqnarray}
